---
title: "hw8_GohBram"
author: "Bram Goh"
format: 
      pdf:
        toc: true
        embed-resources: true
date: "`r Sys.Date()`"
---

```{r}
#| message: false
library(here)
library(readxl)  # for reading excel files
library(modelsummary)  # for summarizing data
library(cmdstanr)  # use two cores
library(posterior)
library(bayesplot)
library(brms)
library(tidyverse)
```

# Research Question

> Do statistical regularities in language production (specifically, content word ratio and combination ratio) predict performance on high predictability SPiN items?

# Variables

- `Highspin_successes`: grouped number of successes on high predictability SPiN trials.
- `Highspin_trials`: number of high predictability SPiN trials
- `content_word_ratio`: average ratio of content words (nouns, verbs, adjective, and adverbs) to total words per sentence
- `combination_ratio`: average ratio of trigrams to total words per sentence

## Data Import

```{r}
frog <- read.csv("processed_frog_data_ver2.csv")
frog[1] <- NULL
frog <- frog %>% select(c(qualtrics_id, content_word_ratio, combination_ratio, Highspin_successes, Highspin_trials))
```

## Variable Summary

Table @tbl-summ-var shows the summary statistics for and Pearson's correlations between the variables of interest.

```{r}
#| label: tbl-summ-var
#| tbl-cap: Descriptive statistics
datasummary_skim(frog)
datasummary_correlation(frog)
```

# Model

Let $Y_i$ = `Highspin_successes`, $N_i$ = `Highspin_trials`, $X_1$ = `content_word_ratio`, $X_2$ = `combination_ratio`

Model:
$$
  \begin{aligned}
    Y_i \sim Bin(N_i, \mu_i)
\\
log(\frac{\mu_i}{1-\mu_i}) = \eta_i
\\
\eta_i = \beta_0 + \beta_1X_1 + \beta_2X_2 + \beta_3X_1 \times X_2
  \end{aligned}
$$

Prior:
$$
  \begin{aligned}
    \beta_0 \sim t_4(0, 1) \\
    \beta_1 \sim t_4(0, 1) \\
    \beta_2 \sim t_4(0, 1) \\
    \beta_3 \sim t_4(0, 1) \\
  \end{aligned}
$$

## Analysis

We used 4 chains, each with 4,000 iterations (first 2,000 as warm-ups). 

```{r}
#| include: false
m_logitlink <- brm(
  Highspin_successes | trials(Highspin_trials) ~ content_word_ratio * combination_ratio,
  family = binomial(link = "logit"),
  data = frog,
  prior = c(prior(student_t(4, 0, 2.5), class = "Intercept"),
            prior(student_t(4, 0, 2.5), class = "b")
           ),
  seed = 1340,
  file = "proj_m1",
  cores = 2,
  chains = 4,
  iter = 4000,
  warmup = 2000)
```

# Results

As shown in the rank histogram in @fig-rank-hist-fit below, the chains mixed well.

```{r}
#| label: fig-rank-hist-fit
#| fig-cap: Rank histogram of the posterior distributions of model parameters.
as_draws(m_logitlink) |>
    mcmc_rank_hist(pars = c("b_Intercept", "b_content_word_ratio", "b_combination_ratio", "b_content_word_ratio:combination_ratio"))
```

```{r}
#| label: tbl-summ-convergence
#| tbl-cap: Posterior summary of the model with convergence statistics.
m_logitlink
```

@tbl-summ-fit shows the summary output for $\beta_0$, $\beta_1$, $\beta_2$, and $\beta_3$, as well as model performance statistics.
```{r}
#| message: false
#| label: tbl-summ-fit
#| tbl-cap: Posterior summary of the model estimates.
msummary(m_logitlink, estimate = "{estimate} [{conf.low}, {conf.high}]",
         statistic = NULL, fmt = 2)
```

```{r}
#| message: false
pp_check(m_logitlink)
```
```

The analysis showed that on average, the content word ratio (M = -2.87, 90% CI [-6.73, 0.57]) and combination ratio (M = 0.96, 90% CI [-1.64, 3.46]) were not significantly associated with performance on high predictability SPiN items. The interaction effect was also non-significant (M = -1.37, 90% CI [-5.39, 2.78]).
